Completing my PhD required the support of numerous people from various chapters of my life, including many people located outside Western Massachusetts.

First and foremost, I thank my advisor, Andrew Cohen. Andrew's support has been crucial to my development as a scientist. Andrew's careful experimental design and thoughtful consideration of all hypotheses are traits that I hope to emulate throughout my career. Andrew was also patient with me beyond words, especially when I was navigating my first few years of graduate school during the pandemic. Moreover, Andrew always has his students' backs and cares for them as people. Andrew is not only a great advisor; he is a kind, thoughtful person whose mentorship enriched my experience as a graduate student. 

Jeff Starns also deserves a tremendous amount of thanks, not only for serving on my dissertation committee but also for his years of mentorship. Jeff has never been my official advisor but has provided a level of support comparable to a thesis chair. Jeff has had tremendous patience for my innumerable questions about Bayesian statistics and math modeling. He has also been a friendly face in the department from day one and was kind enough to invite me to watch a few of his band's gigs. 

I would also like to thank Youngbin Kwak and Elizabeth Miller for serving on my dissertation committee. Their comments and suggestions, throughout the dissertation process, considerably improved the final product. 

Next, I want to thank the incredible friends I have made at UMass, both in and out of the psych department: Anna, Clara, Elijah, Fran, Hector, Jerome, John, Kuan-Jung, Larri, Lisa, Manasa, Mar, Melisa, Michael, Qing, Ramiro, Ryan, Sandarsh, Shimin, Tejas, Tori, Trina, and Yun. In particular, I must thank John and Jerome for being wonderful lab and office mates and for helping me get through COVID. Tejas also deserves a special shout-out for organizing the game nights that helped get me out of my shell at a time I needed it most. Thank you to the rest of the game night crew - Anna, Clara, Elijah, Fran, Manasa, Ramiro, Tori, Trina, and Qing - for a welcome respite from work stress, which I could always look forward to each week. Thank you to Fran for always entertaining me with fun facts about every subject under the sun. Thank you to Melisa for never holding back your opinion on anything. Thank you to Anna for always being a voice of calm and reason. Thank you to Larri for always watching the cats, allowing me the privilege to watch your bunnies, and for being an invaluable friend and colleague. 

I must thank my friends from the Midwest whose support was instrumental to my success, in particular Anton, Ben, Brynn, Emily, Erik, Erin, Gerrit, Kieran, Martin, and Tyler. Thank you all for the love/chaos you bring to my life.

I also want to thank Tim Anderson for providing me with a creative outlet / stress-reduction space by welcoming me into both the UMass Concert Band and Amherst Community Band. 

Without the love and constant support of my family, I could not have made it through graduate school; thank you so so so so much to my parents Pat and Ellen and my siblings Brian, Lyla, and Izzy. Your love has never gone unnoticed or unappreciated. I also want to thank my in-laws, Jim, Ruth, and Ben, for opening your family to me and being understanding, thoughtful, and kind people. 

Lastly, I must thank my wife, Cait. Cait has been my number-one supporter for the entirety of my tenure at UMass. She talked me down from countless moments of panic, supported me while I stressed about my progress in graduate school, and made me laugh every single day, always while reminding me of her love. Not only did she move from Illinois to Massachusetts to be with me during graduate school, she also agreed to drive from Wisconsin to Massachusetts - two days after our wedding - so that I could defend my dissertation. That alone deserves more gratitude than I could possibly ever express. Thank you Cait, from the bottom of my heart.

